Effect of resonant and random excitation on the LHC beam in the context of pulsed hollow electron lens operation.

### 2018-02-13 resex.tex
split section IV in setup and results

\section{LHC experiments on the effect of a resonant and random excitation on the beam core\label{sec:exp}}
%put picture on top
\begin{figure*}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		2016 experiment\\
		\includegraphics[width=0.7\linewidth]{bunchfilling_2016.png}	
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		$ $\\
		2017 experiment\\
		\includegraphics[width=0.7\linewidth]{bunchfilling_2017.png}	
	\end{minipage}
	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom). In 2016 the same excitation amplitude was applied on a group of 4 bunch while in 2016 the number was increased to 6 bunches. In 2016 the excitation was also only applied in one plane leading to $2\times24=48$ bunches in total in 2016 and $3\times2\times36=216$ bunches in 2017.}
\end{figure*}
%put picture on top

@@

\subsection{Simulation and experimental results\label{sec:simex}}
In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations. In summary, the most efficient pulsing patterns could be successfully predicted in simulations, which we will demonstrate on the example of the 2016 experiments (see Sec.~\ref{sec:pattern}). Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8}) as well as the random excitation in Sec.~\ref{sec:simexran}. As a last chapter (Sec.~\ref{sec:damp}), we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.

@@

\subsubsection{Dependence on the pulsing pattern\label{sec:pattern}}
The effect of each pulsing pattern in comparison to the other patterns can be identified by the means of the loss rate and emittance growth from the distribution tracking. As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit1.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance for different pulsing patterns obtained from distribution tracking based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The resonant excitation is applied in both planes with an amplitude of 96~nrad. No random noise component is added.}
\end{figure}

@@

\subsubsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
\begin{enumerate}
	\item Increase of the loss rate with the excitation amplitude (Figs.~\ref{fig:10thexp}.
	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
	\item Change of the beam distribution for the excited bunches (Fig.~\ref{fig:10thexpprof}).	
\end{enumerate}
As loss rates, emittance growth and beam distribution stayed unchanged for the $4$~reference bunches, the above observations can be also truly associated with the effect of the $10^{\mathrm{th}}$~turn pulsing.
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.49\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments. For all bunches the damper was not active. Relative refers here to the normalization to the initial value. For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}). The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation indicated as envelope.}
\end{figure}

@@

\subsubsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the experimental results. Unfortunate for this experiment, the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics considered irrelevant in the context of the presented measurements. The attempt to change back to the fractional tunes of (.28,.31) as chosen in 2016 lead to large losses during the 2017 experiment and therefore the 2017 working point of (62.27,60.295) had to be used instead. This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (see Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (see Fig.~\ref{fig:7th2017fma}).
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
	\end{minipage}	
	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295). The excitation is 96~nrad in both planes. The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
\end{figure}

@@

\subsubsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo, which is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL. With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected from the distribution tracking summarized in Fig.~\ref{fig:8thsim}.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_sigm.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit1.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (distribution tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$. The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
%\end{figure}
%\begin{figure}[t]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$8^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut8skhv_dp0_amp_annotate.png}
	\end{minipage}	
	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for $8^{\mathrm{th}}$~turn pulsing (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295). The excitation is 96~nrad in both planes. The $16Q_y$ and \mbox{$8Q_x-4Q_y$} are both excited.}
\end{figure}

@@

\subsubsection{random excitation\label{sec:simexran}}
The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL. On the other hand this pattern is also the most dangerous one for the proton beam core. The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed. The random excitation is in general so powerful as it excites all frequencies of the beam equally.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		uniform random H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
	\end{minipage}	
	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
\end{figure}
This can be seen formidably on the example of the FMA analysis in amplitude space, depicted in Fig.~\ref{fig:ran2017fmaamp}, for a random excitation as applied during the 2017 experiment. In this case, the diffusion in tune is increased equally by the random excitation independent of the particle amplitude.

@@

\subsubsection{Effect of the transverse damper\label{sec:damp}}
\begin{figure}[b]
	\centering
	random excitation H+V\\
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper not active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
	\end{minipage}	
	\centering
	$7^{\mathrm{th}}$~turn H+V\\
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper not active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
	\end{minipage}	
	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for $7^{\mathrm{th}}$~turn pulsing in V (left) and for a random excitation only in V (right).}
\end{figure}
For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation. On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation. As an example, the vertical emittance growth for the $7^{\mathrm{th}}$~turn pulsing in V  is compared with a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the aforesaid behavior. 
\textcolor{red}{Any idea why? Daniel, help!}

>>>>

\section{Experimental and simulation setup\label{sec:exp}}
%put picture on top
\begin{figure*}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		2016 experiment\\
		\includegraphics[width=0.7\linewidth]{bunchfilling_2016.png}	
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		$ $\\
		2017 experiment\\
		\includegraphics[width=0.7\linewidth]{bunchfilling_2017.png}	
	\end{minipage}
	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom). In 2016 the same excitation amplitude was applied on a group of 4 bunch while in 2016 the number was increased to 6 bunches. In 2016 the excitation was also only applied in one plane leading to $2\times24=48$ bunches in total in 2016 and $3\times2\times36=216$ bunches in 2017.}
\end{figure*}
%put picture on top

@@

\section{Experimental and simulation results\label{sec:simex}}
In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations. In summary, the most efficient pulsing patterns could be successfully predicted in simulations, which we will demonstrate on the example of the 2016 experiments (see Sec.~\ref{sec:pattern}). Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8}) as well as the random excitation in Sec.~\ref{sec:simexran}. As a last chapter (Sec.~\ref{sec:damp}), we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.

@@

\subsection{Dependence on the pulsing pattern\label{sec:pattern}}
The effect of each pulsing pattern in comparison to the other patterns can be identified by the means of the loss rate and emittance growth from the distribution tracking. As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit1.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance for different pulsing patterns obtained from distribution tracking based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The resonant excitation is applied in both planes with an amplitude of 96~nrad. No random noise component is added.}
\end{figure}

@@

\subsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
\begin{enumerate}
	\item Increase of the loss rate with the excitation amplitude (Figs.~\ref{fig:10thexp}.
	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
	\item Change of the beam distribution for the excited bunches (Fig.~\ref{fig:10thexpprof}).	
\end{enumerate}
As loss rates, emittance growth and beam distribution stayed unchanged for the $4$~reference bunches, the above observations can be also truly associated with the effect of the $10^{\mathrm{th}}$~turn pulsing.
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.49\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments. For all bunches the damper was not active. Relative refers here to the normalization to the initial value. For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}). The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation indicated as envelope.}
\end{figure}

@@

\subsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the experimental results. Unfortunate for this experiment, the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics considered irrelevant in the context of the presented measurements. The attempt to change back to the fractional tunes of (.28,.31) as chosen in 2016 lead to large losses during the 2017 experiment and therefore the 2017 working point of (62.27,60.295) had to be used instead. This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (see Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (see Fig.~\ref{fig:7th2017fma}).
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
	\end{minipage}	
	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295). The excitation is 96~nrad in both planes. The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
\end{figure}

@@

\subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo, which is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL. With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected from the distribution tracking summarized in Fig.~\ref{fig:8thsim}.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_sigm.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit1.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (distribution tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$. The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
%\end{figure}
%\begin{figure}[t]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$8^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut8skhv_dp0_amp_annotate.png}
	\end{minipage}	
	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for $8^{\mathrm{th}}$~turn pulsing (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295). The excitation is 96~nrad in both planes. The $16Q_y$ and \mbox{$8Q_x-4Q_y$} are both excited.}
\end{figure}

@@

\subsection{random excitation\label{sec:simexran}}
The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL. On the other hand this pattern is also the most dangerous one for the proton beam core. The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed. The random excitation is in general so powerful as it excites all frequencies of the beam equally.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		uniform random H+V
		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
	\end{minipage}	
	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
\end{figure}
This can be seen formidably on the example of the FMA analysis in amplitude space, depicted in Fig.~\ref{fig:ran2017fmaamp}, for a random excitation as applied during the 2017 experiment. In this case, the diffusion in tune is increased equally by the random excitation independent of the particle amplitude.

@@

\subsection{Effect of the transverse damper\label{sec:damp}}
\begin{figure}[b]
	\centering
	random excitation H+V\\
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper not active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
	\end{minipage}	
	\centering
	$7^{\mathrm{th}}$~turn H+V\\
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper not active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		damper active
		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
	\end{minipage}	
	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for $7^{\mathrm{th}}$~turn pulsing in V (left) and for a random excitation only in V (right).}
\end{figure}
For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation. On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation. As an example, the vertical emittance growth for the $7^{\mathrm{th}}$~turn pulsing in V  is compared with a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the aforesaid behavior. 
\textcolor{red}{Any idea why? Daniel, help!}
